---
title: "ARIMA_testing"
author: "Ben Wortman"
date: "4/27/2020"
output: pdf_document
---

# to do:
# Fix the Y label for the graphs
# ACF <- auto correlation function 
# IQR of the original data, and IQR of the final Residuals
# SNR of original vs SNR of the Final
# Toy examples 
# Look at some real stars
# Graph some of davenports stars

# plot for the 70k points
# plot(tso_resid1[1:80], pch=20, cex=0.5, col="#11111130",main="Residuals")

# acf(tso_resid1, xrange=c(1:30))
# SNR <- max(tso_resid1) / IQR(tso_resid1)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("changepoint")
library("tsoutliers")
library("dtw")
library("lubridate")
library("symmetry")
```

## Synthesize The Data
First set up the testing sequence we want our autoarima model to latch onto

```{r}
# generate a time series to run arima on
seq_length <- 1000 # how long of a sequence to generate
star_rotation_hours <- 24 # the synthetic stars rotation time
noise <- 1 # how active the synthetic star is 
star_amplitude <- 5
# generate a 10k length sinosoidal sequence for the stars rotation and amplitude
sequence <- star_amplitude * sin(seq(0,2*pi, 2*pi / star_rotation_hours / 4))
sequence <- rep(sequence,seq_length / round(length(sequence)))
# add in the random noise
sequence_raw <- sequence + rnorm(length(sequence), 0, noise)

```


## Generate the Flare Shape
Next we approximate the shape of a flare using the exponential function and then insert it randomly into the synthetic data. 

```{r}
# generate the flare shape - typically 2 hour decay
intensity <- 10
time_periods <- 8
exp_outlier <- intensity * dexp(seq(0.5,20,0.25),2)[1:time_periods] / dexp(0.5,2)
#plot(exp_outlier)

# set the flare position randomly in the middle 50 percent of the data 
position <- round(runif(1,round(seq_length * .25),round(seq_length * 0.75)))

# duplicate the original residuals 
sequence_flare <- sequence_raw

# insert the flare into the assigned position
sequence_flare[position:(position+length(exp_outlier)-1)] <- sequence_flare[position:(position+length(exp_outlier)-1)] + exp_outlier



# get the parameters for the outliers library
#pars <- coefs2poly(fit)
# test tsoutliers ovre the residuals
#tso_outliers <- locate.outliers(resid, pars, cval=3.5, types = c("TC","AO"), delta = 0.5)
#tso_outliers

```
## Fit to ARIMA 
Then proceed to fit the ARIMA model to the raw data with the injected flare. 
```{r}
# fit using arima
fit <- auto.arima(sequence_flare, max.order = 10, stepwise = F)

# Get the residuals from arima
resid <- residuals(fit)
```


## Plot the Results 
```{r}
# plot the residuals example
par(mfrow=c(2,2))
plot(c(rep(0,40), exp_outlier, rep(0,40 - time_periods)), type = "l", main="Flare Shape", ylab="Intensity")
# raw data 
plot(sequence_raw[(position-40):(position+40)], type = "l",main="Raw Data (No Flare)", ylab="Intensity")
# raw with flare
plot(sequence_flare[(position-40):(position+40)], type = "l", main="Raw Data (With Flare)", ylab="Intensity")
abline(v=40, lty=2, col="red")
# resids
plot(resid[(position-40):(position+40)], type = "l", main="Residuals", ylab="Residuals")
abline(v=40, lty=2, col="red")

# Signal to Noise Ratio -> meaningful noise / IQR
SNR <- max(resid) / IQR(resid)
```

