---
title: "Synthetic_ARIMA_flare_profiles"
author: "Ben Wortman"
date: "4/27/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Generating Synthetic Flare Profiles:
By repeating the steps in ARIMA_testing_markdown.Rmd for hundreds of flares and then overlaying the resulting flare profiles, we can get a sense of what these flare profiles look like after being passed through ARIMA. We generated 100 flares per 20000 timestamps in the sequence to speed up the sampling. 

```{r}
library("changepoint")
library("tsoutliers")
library("dtw")
library("lubridate")
library("forecast")

generate_profiles <- function(flare_intensity, 
                              star_noise=1, 
                              buffer_periods=2,
                              seq_length = 20000,
                              flare_num = 100){

  # generate a time series to run arima on
  seq_length <- seq_length # how long of a sequence to generate
  star_rotation_hours <- 24 # the synthetic stars rotation time
  noise <- star_noise # how active the synthetic star is 
  star_amplitude <- 5
  # generate a 10k length sinosoidal sequence for the stars rotation and amplitude
  sequence <- star_amplitude * sin(seq(0,2*pi, 2*pi / star_rotation_hours / 4))
  sequence <- rep(sequence,seq_length / round(length(sequence)))
  # add in the random noise
  sequence_raw <- sequence + rnorm(length(sequence), 0, noise)
  
  # generate the flare shape - typically 2 hour decay
  intensity <- flare_intensity
  time_periods <- 8
  exp_outlier <- intensity * dexp(seq(0.5,20,0.25),2)[1:time_periods] / dexp(0.5,2)
  #plot(exp_outlier)
  
  # set the flare position randomly in the middle 90 percent of the data 
  flare_num <- flare_num # the number of flares to be inserted into the sequence
  position <- round(runif(flare_num,round(seq_length * .05),round(seq_length * 0.95)))
  
  # duplicate the original residuals 
  sequence_flare <- sequence_raw
  
  # insert the flare into the assigned position
  for(pos in position){
  sequence_flare[pos:((pos+length(exp_outlier)-1))] <- sequence_flare[pos:((pos+length(exp_outlier)-1))] + exp_outlier}
  
  # fit using arima
  fit <- auto.arima(sequence_flare, max.order = 10, stepwise = F)
  
  # Get the residuals from arima
  resid <- residuals(fit)
  
  
  # get the star locations into a matrix
  flare_position_label <- sprintf("t(%s)", seq(-buffer_periods,time_periods+buffer_periods+1))
  flare_point_matrix <- c()
  for(i in -buffer_periods:(time_periods+buffer_periods+1)){
    flare_point_matrix <- cbind(flare_point_matrix, position + i)
  }
  flare_point_matrix
  
  # get the residuals into a matrix to be returned. Each column is a time
  residual_matrix <- c()
  for(i in 1:dim(flare_point_matrix)[2]){
    residual_matrix <- cbind(residual_matrix, resid[flare_point_matrix[,i]])
  }
  residual_matrix <- data.frame(residual_matrix)
  names(residual_matrix) <- flare_position_label
  residual_matrix
}

```


# Plot the Profile
Next we can generate and then plot the profiles for all the synthesized data. 
```{r}
# create a function to plot the flare profiles
point_plot <- function(flare_profile, intensity, lower=-3, upper=5){
  # create an empty plot to add the points to
  plot(1, type="n", 
       axes=T, 
       xlab="Time of Event", 
       ylab="Intesity",
       xlim=c(-3,11), 
       ylim=c(lower,upper), 
       main=paste0("Flare ",intensity,"x the intensity of noise"))
  # Plot the points using a for loop
  for (i in 1:(dim(flare_profile)[2])){
    points(rep(i-3, dim(flare_profile)[1]), flare_profile[,i], pch=20, cex=0.5, col="#11111130")
  }
}

# create a function to get the flare profiles quicker
get_profiles <- function(intensity){
  # generate the matrix of flare profiles
  profiles <- c()
  # get 1000 profiles
  for (i in 1:10){profiles <- rbind(profiles, (generate_profiles(intensity)))}
  profiles
  
}

# write a function to save the flare profile
save_profile <- function(flare_matrix, intensity_label){
  filename <- paste("../synth_profiles/", intensity_label,"x_initial_signal_to_noise.csv", sep = "")
  write.csv(flare_matrix, filename, row.names = F)
}


# loop through different intensity values and save the results
# star and flare parameters can be adjusted in the main function above
for (i in 8:9){
  profiles <- get_profiles(i)
  save_profile(profiles,i)
}

# then load the csvs to look at the estimated flare profiles after arima
# 2x flare intensity
profile <- read.csv("../synth_profiles/2x_initial_signal_to_noise.csv")
point_plot(profile,2,-4,5)

# 3x flare intensity
profile <- read.csv("../synth_profiles/3x_initial_signal_to_noise.csv")
point_plot(profile,3,-4,6)
boxplot(profile)

# 4x flare intensity
profile <- read.csv("../synth_profiles/4x_initial_signal_to_noise.csv")
point_plot(profile,4,-4,7)
boxplot(profile)

# 5x flare intensity
profile <- read.csv("../synth_profiles/5x_initial_signal_to_noise.csv")
point_plot(profile,5,-4,8)
boxplot(profile)

# 6x flare intensity
profile <- read.csv("../synth_profiles/6x_initial_signal_to_noise.csv")
point_plot(profile,6,-4,9)
boxplot(profile)

# 7x flare intensity
profile <- read.csv("../synth_profiles/7x_initial_signal_to_noise.csv")
point_plot(profile,7,-4,12)
boxplot(profile)



```

## SNR for the different intensities
We calculated the SNR using mean peak signal for the simulated flares divided by the IQR of 10 dummy lightcurves.

```{r}


# Get the IQR for the artificial data by averaging the IQR for 30 sequences
iqr_list <- c()
var_list <- c()
for(i in 1:30){
  # generate a time series to run arima on
  seq_length <- 10000 # how long of a sequence to generate
  star_rotation_hours <- 24 # the synthetic stars rotation time
  noise <- 1 # how active the synthetic star is 
  star_amplitude <- 5
  # generate a 10k length sinosoidal sequence for the stars rotation and amplitude
  sequence <- star_amplitude * sin(seq(0,2*pi, 2*pi / star_rotation_hours / 4))
  sequence <- rep(sequence,seq_length / round(length(sequence)))
  # add in the random noise
  sequence_raw <- sequence + rnorm(length(sequence), 0, noise)
  
  # fit using arima
  fit <- auto.arima(sequence_raw, max.order = 10, stepwise = F)
  
  # Get the residuals from arima
  resid <- residuals(fit)
  
  # add the IQR to the list
  iqr_list[i] <- IQR(resid)
  var_list[i] <- var(resid)
  print(i)
}

# get the mean IQR
mean_variance <- mean(var_list)
mean_IQR <- mean(iqr_list) # 1.6
SNR_list <- c()
for(file in list.files("../synth_profiles")){
  profile <- read.csv(paste0("../synth_profiles/",file))
  # get the SNR
  SNR_list <- c(SNR_list,  mean(profile$t.0.) / mean_IQR)
  print(c(file, mean(profile$t.0.) / mean_IQR, mean(profile$t.0.) / mean_variance))
}



# get the IQR from the average of 10 ARIMA models 



# Signal to Noise Ratio -> meaningful noise / IQR
SNR <- max(resid) / IQR(resid)
```

Plots for the synthetic flare profiles 
```{r}
par(mfrow=c(2,2))
# then load the csvs to look at the estimated flare profiles after arima
# 2x flare intensity
profile <- read.csv("../synth_profiles/2x_initial_signal_to_noise.csv")
point_plot(profile,2,-4,5)

# 3x flare intensity
profile <- read.csv("../synth_profiles/3x_initial_signal_to_noise.csv")
point_plot(profile,3,-4,6)

# 4x flare intensity
profile <- read.csv("../synth_profiles/4x_initial_signal_to_noise.csv")
point_plot(profile,4,-4,7)

# 5x flare intensity
profile <- read.csv("../synth_profiles/5x_initial_signal_to_noise.csv")
point_plot(profile,5,-4,8)

# 6x flare intensity
profile <- read.csv("../synth_profiles/6x_initial_signal_to_noise.csv")
point_plot(profile,6,-4,9)

# 7x flare intensity
profile <- read.csv("../synth_profiles/7x_initial_signal_to_noise.csv")
point_plot(profile,7,-4,10)

# 8x flare intensity
profile <- read.csv("../synth_profiles/8x_initial_signal_to_noise.csv")
point_plot(profile,8,-4,11)

# 10x flare intensity
profile <- read.csv("../synth_profiles/10x_initial_signal_to_noise.csv")
point_plot(profile,10,-4,13)



```


## Checking the Bias of the synthetic data with flares inserted

```{r}

# generate a time series to run arima on
flare_intensity <- 8
flare_num = 300
seq_length <- 70000 # how long of a sequence to generate
star_rotation_hours <- 24 # the synthetic stars rotation time
noise <- 1 # how active the synthetic star is 
star_amplitude <- 5

# run the test for different intensities
bias_values <- c()
time <- Sys.time() 
for (flare_intensity  in 2:10){
  # generate a 10k length sinosoidal sequence for the stars rotation and amplitude
  sequence <- star_amplitude * sin(seq(0,2*pi, 2*pi / star_rotation_hours / 4))
  sequence <- rep(sequence,seq_length / round(length(sequence)))
  # add in the random noise
  sequence_raw <- sequence + rnorm(length(sequence), 0, noise)
  
  # generate the flare shape - typically 2 hour decay
  intensity <- flare_intensity
  time_periods <- 8
  exp_outlier <- intensity * dexp(seq(0.5,20,0.25),2)[1:time_periods] / dexp(0.5,2)
  #plot(exp_outlier)
  
  # set the flare position randomly in the middle 90 percent of the data 
  flare_num <- flare_num # the number of flares to be inserted into the sequence
  position <- round(runif(flare_num,round(seq_length * .05),round(seq_length * 0.95)))
  
  # duplicate the original residuals 
  sequence_flare <- sequence_raw
  
  # insert the flare into the assigned position
  for(pos in position){
  sequence_flare[pos:((pos+length(exp_outlier)-1))] <- sequence_flare[pos:((pos+length(exp_outlier)-1))] + exp_outlier}
  
  # fit using arima
  fit <- auto.arima(sequence_flare, max.order = 10, stepwise = F)
  
  # Get the residuals from arima
  resid <- residuals(fit)
  
  # get the parameters for the outlier library
  pars <- coefs2poly(fit)
  
  tso_outliers <- locate.outliers(resid, pars, cval=3.5, types = c("AO"), delta = 0.5)
  #tso_outliers
  
  # get the bias of the outlier distribution
  bias_values <- c(bias_values, length(tso_outliers$ind[tso_outliers$coefhat > 0]) - length(tso_outliers$ind[tso_outliers$coefhat < 0]))
  print(time - Sys.time())
  time <-Sys.time() 
}

plot(2:10, bias_values, ylab="Bias", xlab="Flare Intensity", main="Bias of 300 Synthetic Flares")

```

